---
title: "Module 1: Introduction to R"
tutorial:
  id: "module1"
  name: "Module 1"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Basic Introduction to `R` Environment and Tutorial Access"
---

```{r setup, include=FALSE}
library(learnr)
library(tutorial.helpers)
options(repos = c(CRAN = "https://cran.rstudio.com"))
```

```{r info-section, child=system.file("child_documents/info_section.Rmd", package = "tutorial.helpers"),echo=F}
```

```{r copy-code-chunk, child=system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

## Introduction

In this tutorial, we will introduce some of the basics regarding `R` code and the `R` environment. We will also explain how to access and run tutorials for this class, as well as externally created tutorials that accompany the text *R for Data Science (2e)*. As a reminder, the book is available at this link [here](https://r4ds.hadley.nz).

`R` certainly has a learning curve to it, and it can be more complicated than other programs. This is because it is also very flexible and adaptable. `R` can accomplish complex and tedious computations in a few lines of code, but the tradeoff is that our code must be very precise. 

In the `R_Setup` tutorial, we learned about the general form of functions:

```{r,function1, eval=F}
functionName(argument1 = value1, argument2 = value2, and so on)
```

and we learned two very important functions: `install.packages()` and `library()`.

### Exercise 1

To reinforce what we've learned, let's load the `tidyverse` package below.

```{r load-library, exercise=TRUE}

```

```{r load-library-hint-1, eval = FALSE}
library(...)
```

```{r load-library-test, include = FALSE}
library(tidyverse)
```

###

We will install and use many different packages throughout the semester. Let's install a new one now called `r4ds.tutorials`. Complete the text below: 

```{r install-package, exercise=TRUE}

```

```{r install-package-hint-1, eval = FALSE}
install.packages("pkg_name_must_be_in_quotes")
```

```{r install-package-test, include = FALSE}
install.packages("r4ds.tutorials")
```

###

We will rely on this package extensively, especially in the early stages of the course. This package was written and recently distributed by David Kane:

Kane D (2025). r4ds.tutorials: Tutorials for "R for Data Science". R package version 0.3.2.9003, https://ppbds.github.io/r4ds.tutorials/.

## `R` and RStudio Basics

Rstudio includes 4 windows:

- **The console** is where code is executed. Here, input and output are shown, and code can be written and executed directly
- **script** or syntax editor. Here we write longer lines of code and then send all or part of the script to the console to run it.
- The RStudio **Environment** which shows available data, saved objects, functions, script history, and **includes a tutorial tab**
    - Environment also includes `Import Dataset` button, a point-and-click convenience function for reading in data
- **utility** pane, which can show plots, tables, help windows, and file paths

### Objects

When performing data analysis in `R`, we often need to create longer series of codes with multiple types of functions and multiple things (e.g., variables) that we perform those functions on. Thankfully, `R` is an object-oriented coding language, meaning we can create placeholders called **"objects"** that store data, functions, or even the result of functions. 

###

To create an object, we use the *assignment* operator, `<-`. The assignments have the form of 

```{r, eval=F}
object <- value
```

We often read the assignment operator as "gets", as in, "the `object` **gets** this `value`.

###

Here is an example. In the code chunk below, type `8 + 5` and run code. This is equivalent to running this math once in the `R` console. 

```{r math1, exercise=TRUE}

```


```{r math1-hint-1, eval=F}
8 + 5
```


###

The result of the function (the `+` is a math function) is printed in the console; however, if we wanted to do the calculation a second time, we'd have to re-type `8 + 5`. Instead, we can assign the number 13 to an object:

```{r, object1, exercise=T}
thirteen <- 13
```

(be sure to hit "Run Code")

Now if we call this object in the console or by running code, it will output what is stored in the object - namely, the number 13.

```{r, object2, exercise=T}
thirteen
```
###

We can also assign the result of a function, such as `8 + 5`, to an object. 


```{r math2, exercise=TRUE}
thirteen <- ...
thirteen
```

To make a short story long, what's happening above is that `R` recognizes `8` and `5` as `numeric` objects, and the `+` is a function that performs addition to numeric objects. The result is the number thirteen, which is saved in the assigned object `thirteen`. This is why it does not simply print out `8 + 5`. 
###

`R` is very sensitive to the type of variable or input that it is receiving. The most common types are `numeric` (also called `double` or `dbl`), `factor` or `fct`, and `character` or `chr`. For example, we can tell `R` that the variable is a character by using quotations. Run the following to see the difference: 

```{r math3, exercise=TRUE}
thirteen <- "8 + 5"
thirteen
```

Our reading assignment this week discusses other types of data in `R`. We will cover variable types and some of these coding quirks extensively throughout the semester. 

## Accessing Tutorials

As mentioned, we will use tutorial modules like this one as the bulk of our content throughout the semester. There are a suite of pre-built tutorials in the `r4ds.tutorials` package and the `learnr` package. Additionally, I will create tutorials that can be accessed through a custom `R` package shared through GitHub. First, let's install some packages needed to access all of these sources. 

Be sure to type or copy/paste this code onto your local RStudio console and make sure it runs. 

```{r, eval=F}
install.packages("learnr")
install.packages("remotes")
install.packages("tutorial.helpers")
#install.packages(r4ds.tutorials) already completed previously
```

Typically, the tutorial scripts will also load any necessary packages, but it is useful to have all packages installed locally, just in case. 

###

To launch a tutorial on your own computer (laptop or lab PC), we will use a function from `learnr` that is called `run_tutorial()`. In this function, you will provide the name of the tutorial and the name of associated package. Here is an example: 

```{r, eval=F}
learnr::run_tutorial("rstudio", package = "tutorial.helpers")
```

As an aside, this specific syntax above with the double colon `::` is a special use where we call a function from a specific package without loading the entirety of the package into the active library. In other words, `learnr` is installed, but you do not have to run `library(learnr)` to access the `run_tutorial()` function when it is written this way. 

###

The `rstudio` tutorial above is the next module that you will go through. After `learnr` and `tutorial.helpers` are installed, you can run that line above to launch the tutorial in an RStudio/web browser window. 
Each time a tutorial is assigned, I will provide the one-liner code needed to open it. 

### `r4psych.tutorials`

I have also created a custom package containing tutorials, exercises, and homework assignments for this class. To access tutorials for additional content or assignments in our course, you will need to download my package from GitHub. The package is called `r4psych.tutorials`. Always install the latest version before launching a new tutorial or assignment module. To install, run: 

```{r, eval=F}
remotes::install_github("luningham-lab/r4psych.tutorials")
```

To walk through this: 
- `remotes` is an additional package that can access content from the Web. 
- `install_github` is a function that will install `R` packages that are uploaded to the website GitHub and not necessarily available on the Comprehensive `R` Archive Network (CRAN). Packages can be updated and distributed much quicker through GitHub than CRAN. 
- `luningham-lab` is my GitHub account, and this directs the `install` function to the correct user's library to find the package.

### 

This package is a living and breathing entity, so it will be important that you re-install it with the above code frequently.

Now you can load tutorials from this package (again, code will be provided):

```{r, eval=F}
learnr::run_tutorial("module1", package = "r4psych.tutorials")
```

## Knowledge Check 

Suppose we need to use a new package for our research, such as the `psych` package. In the code chunk below, write both lines of code needed to install the package and then load it into the active library. 

```{r knowledgecheck1, exercise=TRUE}


```

### Question 2

In the RStudio IDE, what is the difference between the syntax editor window and the console window?

```{r knowledgecheck2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

### Question 3

What is the purpose/usage of the assignment operator, and what is its symbol?

```{r knowledgecheck3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

### Question 4

Below, assign a number to the object `x1` and another number to the object `x2`. Then, assign their sum to the object `y`. Examine `y` (run it on its own line) to make sure everything executed as expected.  

```{r knowledgecheck4, exercise=TRUE}


```

### Question 5

```{r quiz1, echo=FALSE}
quiz(caption = "Tutorials and GitHub Packages",
  question("What function will load a tutorial from an installed package?",
     answer("library(tutorial_name)", message = "No, the tutorial itself is not a standalone package"),
    answer("tutorials(tutorial_name, package name"),
    
    answer("learnr::run.Tutorials(\"tutorial_name\", package=\"package\" ", message="not quite - R is very particular about exact symbols and letter case."),
    answer("learnr::run_tutorial(\"tutorial_name\", package=\"package\" ", correct=T),
    allow_retry = TRUE
  ),
  question("How can we install/update the class package from GitHub?",
answer("library(r4psych.tutorials)", message = "This will load the previously downloaded version, but it won't install or update the cotents of the package."),
    answer("GitHub::install.packages(\"r4psch.tutorials\")"),
    answer("remotes::install.packages(\"r4psch.tutorials\")", message="not quite - install.packages is from base R, not from the remotes package, and we need to tell R the specific user's library to find the package."),
    answer("remotes::install_github(\"luningham-lab/r4psych.tutorials\")", correct=T),
    allow_retry = TRUE
  )
)
```



```{r download-answers, child=system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers"),echo=F}
```
