---
title: "Binomial regression in the generalized lienar model"
tutorial:
  id: "module14-binomial"
  name: "Binomial regression in R"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
author: Justin Luningham
description: "Binomial regression in R."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(MASS)
library(car)
library(emmeans)
library(afex)
library(kableExtra)
library(jtools)
library(infer)
library(moderndive)
library(learnr)
library(report)
library(rempsyc)
library(broom)
library(data.table)
library(flextable)
library(rstatix)
library(ggpubr)
library(tidyverse)
library(tutorial.helpers)
library(datarium)
library(GGally)
library(mosaicData)
options(repos = c(CRAN = "https://cran.rstudio.com"))
if (!requireNamespace("gradethis", quietly = TRUE)) {
  learnr::tutorial_warning(
    "This tutorial uses the **gradethis** package to provide feedback.  
    It looks like you don’t have gradethis installed yet.  
    Please run this in your console:
    
    remotes::install_github('rstudio-education/gradethis')"
  )
} else {
  library(gradethis)
  gradethis::gradethis_setup()  # optional: standardizes grading defaults
}

set.seed(144445)
one_sample <- tibble(first_walk = round(rnorm(50, 12.8, 2.5),1))
raw_pairs<-mvrnorm(25, mu = c(14, 12.8), Sigma = cbind(c(2.5, 1), c(1, 2.5)))

paired_samples <- tibble(sib1 = round(raw_pairs[,1],1), sib2 = round(raw_pairs[,2],1), famID=1:25)

paired_long <-paired_samples |> pivot_longer(cols = c(sib1, sib2), values_to = "first_walk", names_to = "sibling")

two_sample<-tibble(control = round(rnorm(50, 14, 2.5),1), intervention = round(rnorm(50, 13, 2.5 ),1))

two_sample_long <- two_sample |> 
  pivot_longer(cols = c(control, intervention), # or, cols = everything()
              values_to = "first_walk", names_to = "group")

#dat <- read_delim("~/Downloads/Tab10-2.txt", 
#    delim = "\t", escape_double = FALSE, 
#    trim_ws = TRUE)

attr(stress$score, "label") <- "Stress Score"
attr(stress$age, "label") <- "Age"

#contrasts(stress$treatment) <- contr.sum(2)
#contrasts(stress$exercise) <- contr.sum(3)





fit2 <- glm(outcome ~ age, Whickham, family = "binomial")
effect_plot(fit2, pred = "age", interval = TRUE, plot.points = TRUE)
effect_plot(fit2, pred = "age", interval = TRUE)
summary(fit2)

fit3 <- glm(outcome ~ smoker, Whickham, family = "binomial")
summary(fit3)

effect_plot(fit3, pred = "smoker")

fit4 <- glm(outcome ~ smoker + age, Whickham, family = "binomial")
summary(fit4)

effect_plot(fit4, pred="age", interval=T)
effect_plot(fit4, pred="age", interval=T, partial.residuals = T)



#fit5 <- glm(outcome ~ smoker + age_cat, Whickham, family = "binomial")

```


```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Categorical outcomes
###

All of the models we've learned so far assume that the outcome is a continuous, quantitative variable with errors that have a mean of 0 and a constant variance. In many cases, we also assume that these errors are normally distributed, which means that $Y$ conditioned on $X$ is also normally distributed. 

Of course, many outcomes that we study are not exactly continuous and are certainly not normally distributed. We mentioned that many of these approaches (e.g. t-test, ANOVA, even linear regression) are **robust** to some violations of assumptions. They are still accurate even if data are not perfectly normal or variances are not equal across groups. 

Will this hold if the outcome is not truly continuous? 

###

As usual, the answer is "it depends!" How different from continuous is our outcome? 

If the outcome is a Likert-type variable on 1-7 scale with $\sim$ 500 responses, it might be approximately normal. But what if the outcome is a specific **event**, such as answering a question correctly or receiving a diagnosis?

###

Let's revisit the `Whickham` dataset examining smoking status and mortality among Women in Whickham, UK. The dataset is a subset of women who were current smokers or never smokers between 1972-1974, and their survival status was ascertained after 20 years. The variables are `outcome`, `smoker`, and `age` at the time of first survey. 

```{r fit1, exercise = TRUE}
library(mosaicData)
glimpse(Whickham)
```

###

Fit a simple linear model predicting `outcome` from `age` below. Call it `fit1`. 

```{r fit2, exercise = TRUE}
fit1 <- ___(___, Whickham)
```

###

The warning message occurs because `lm` struggles internally when the outcome is a factor, not a number. Let's quickly create a new column called `outcome2`, which converts `outcome` to a numeric value. 

```{r fit3, exercise = TRUE}
Whickham <- Whickham |>
  ___(outcome2 = (as.numeric(___) - 1))
Whickham |> count(outcome, outcome2) #check
```


```{r fit3-2, include = FALSE}
Whickham <- Whickham |> mutate(outcome2 = as.numeric(outcome)-1)
```

I include the `-1` part here to scale `outcome2` as 0 or 1. 

###

Now, re-run `fit1` with `outcome2` as the outcome. Check the `effect_plot` with `age` as the predictor. Be sure to plot the raw data points.  

```{r fit4, exercise = TRUE, exercise.setup = "fit3-2"}
fit1 <- lm(outcome ~ age, Whickham)

___(fit1)
effect_plot(fit1, pred = ___, 
            interval = ___, 
            plot.points = ___)
```

```{r fit4-2, include = F}
Whickham <- Whickham |> mutate(outcome2 = as.numeric(outcome)-1)
fit1 <- lm(outcome2 ~ age, Whickham)
```

###

Let's examine the residuals against the fitted values. `augment()` the fit and pipe it to a simple `ggplot` for residuals on the `y` axis and fitted values on the `x` axis. 

```{r fit5, exercise = TRUE, exercise.setup = "fit4-2"}
___(fit1) |> 
  ggplot(aes(x = .fitted, y = .resid))+
  geom____()
```

###

Yikes!

### The generalized linear model 

Clearly, there are instances where a linear predictor of a continuous $Y$ simply does not make sense. Statistically, it's a mess, and practically, it's impossible to interpret. An "event" outcome is a clear example. 
With a binary outcome, the observed data are simply zeros and ones, and they follow a specific distribution called a Bernoulli distribution. This is a special case of the binomial distribution with only one "trial". The parameter of the Bernoulli distribution is $p$, of the probability that the outcome equals 1 (probability of "success"). 

###

For example, a single coin flip follows a Bernoulli distribution with $p = 0.5$. If 1 = Heads, then $P(Y = Heads)$ = 0.5 and $P(Y = Tails) = 1 -P(Y = Heads) = 0.5$. 

In our regression model, we don't want to model the 0's and 1's directly; instead, we explicity model $P(Y = 1)$, or $p$. 

###


The **generalized** linear model extends the general linear model by applying some "link function" to the outcome, such that it is transformed in a way that corresponds to the unique distribution of the data. 

$$ g(Y) = \beta_0 + \beta_1 X_{1} + \beta_2 X_{2} +... +\beta_k X_k$$ 
The key take-aways are that

1. The outcome is still "linear in the predictors" - effects are additive
2. However, we have to interpret the regression coefficients as a function of the *inverse* of the link function
3. The errors/residuals are not included because they look different depending on the form of the link 

###

For binary outcomes, the probability that $Y = 1$, or $p$, is linked to the outcomes through the log-odds link:

$$ \log (\frac{p}{1-p}) = \beta_0 + \beta_1 X_{1} + \beta_2 X_{2} +... +\beta_k X_k$$
This outcome is also called the "log odds".

###

This all happens internally with our statistical software; however, the estimates are usually on the un-transformed scale. The inverse of the log odds is to take the exponential of everything on the right-hand side. 

###

Enough of that, let's see what this means in practice. The base `R` function for the generalized linear model is conveniently called `glm()`. It actually operates a lot like the `lm`, but it requires an additional argument called `family`. The `family` refers to the type of distribution of the outcome. For 0/1 data, we add `family = "binomial". 

The following fits the model and gets the summary, in the same way as `lm` - with slightly different information.

```{r fit6, exercise = TRUE, exercise.setup = "fit4-2"}
fit2 <- glm(outcome ~ age, Whickham, 
            family = "binomial")
summary(fit2)
```

###

In `fit1`, the age effect was 0.017, or for every year increase, the predicted increase on 0-1 mortality was 0.017. Now the age estimate is 0.12. What does this mean? 

Recall that this estimate is on the untransformed scale. One year increase in age corresponds to 0.12 increase in the log-odds of death. For these coefficients, a positive value indicates an increase in the predictor leads to an increase likelihood of the outcome, whereas a negative coefficient suggests a decrease in the likelihood.

If we take the inverse of the log, we get something much more interpretable: the odds ratio. This is the multiplicative change in the odds of the outcome per one-unit change in the predictor. No effect is indicated by odds ratio of 1.0. Greater than 1 indicates an increase in odds, less than one indicates a decrease in the odds.

###

The inverse is found by "exponentiating" the regression coefficients. We can do this in `R` with `exp()`. 

More conveniently, we can `tidy` our output and ask for the transformation internally. 

```{r fit7, exercise = TRUE, exercise.setup = "fit6"}
tidy(fit2, exponentiate =  TRUE, conf.int = TRUE)
```

This means that every one-year increase in age leads to a 13% increase in the odds of death. 

```{r fit8, exercise = TRUE, exercise.setup = "fit6"}
effect_plot(fit2, pred = "age", interval = TRUE)
```

###

[under construction]

- add smoker status
- deviance and LRT 
- predictions

```{r cat_age, include = F}
Whickham<- Whickham |>
  mutate(age_cat = case_when(
    age <= 44 ~ "18-44",
    age > 44 & age <= 64 ~ "45-64",
    age > 64 ~ "65+"))
```

###

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
