---
title: "One-Way Contrasts in R"
subtitle: "PSYC 50523"
author: "Justin Luningham"
tutorial:
  id: "ANOVA_one-way_contrasts"
  name: "One-Way Contrasts"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Pairwise and complex comparisons using aov() and emmeans"
---

```{r setup, include=FALSE}
library(learnr)
library(knitr)
library(kableExtra)
library(dplyr)
library(AMCP)
library(emmeans)
options(max.print = 40)

if (requireNamespace("gradethis", quietly = TRUE)) {
  library(gradethis)
  gradethis::gradethis_setup()
}

knitr::opts_chunk$set(echo = TRUE, comment = "", warning = FALSE, message = FALSE)
```

## Welcome {data-progressive=FALSE}

In this tutorial, we will build on what we learned in the Introduction to R module. We will:

- Review the one-way ANOVA using `aov()`
- Introduce the `emmeans` package for **estimated marginal means**
- Conduct **pairwise comparisons** as follow-up tests after a significant omnibus ANOVA
- Set up and test **complex (custom) contrasts**

This tutorial assumes you are comfortable with loading packages, using `dplyr` and the pipe operator, converting variables to factors, and running `aov()`. If any of that is unfamiliar, revisit the Introduction to R tutorial first.

###

**Estimated time:** ~30 minutes

## A Motivating Example

### The Hypertension Data

Table 4.1 from the text presents hypothetical data for four groups of participants receiving different hypertension treatments, measured on systolic blood pressure. The four treatment groups are: Drug Therapy, Biofeedback, Diet, and Combination.

Let's start by loading and preparing the data. We will use `case_match()` from `dplyr` to recode the numeric condition variable into meaningful labels. Recall from the first tutorial that `mutate()` adds or modifies columns in a data frame, and that grouping variables must be converted to **factors** for ANOVA.


```{r data1, exercise=TRUE}
library(dplyr)
library(AMCP)

data(C4T1)

head(C4T1)
str(C4T1)
```

###

Now let's recode the conditions and convert to factors. The `case_match()` function takes the variable to recode, then a series of formulas: `old_value ~ "new_label"`. We wrap this inside `mutate()` to create the new variable:

```{r data2, exercise=TRUE}
C4T1 <- C4T1 |> 
  mutate(group = case_match(cond,
    1 ~ "Drug Therapy",
    2 ~ "Biofeedback",
    3 ~ "Diet",
    4 ~ "Combination"
  )) |>
  mutate(group = factor(group), cond = factor(cond))

str(C4T1)
```

Notice the use of the **native pipe** `|>`. This works the same as the `%>%` pipe from the first tutorial --- it passes the result on the left into the first argument of the function on the right. The `|>` pipe is built into base R (version 4.1+), so it does not require any additional packages.

### Group Summary Statistics

Let's recreate the summary table from the text, as we practiced in the first tutorial with `group_by()` and `summarize()`:

```{r data3, exercise=TRUE,exercise.setup="data2"}
Summary.4.1 <- C4T1 |> 
  group_by(group) |>
  summarize(
    Mean = mean(bloodpr),
    Median = median(bloodpr),
    Var = var(bloodpr),
    SD = sd(bloodpr),
    n = n()
  )

Summary.4.1
```

### Exercise: Visualize the Data

Before running any statistical tests, create a boxplot of `bloodpr` by `group` to visually inspect the distributions. Use either base R `boxplot()` or `ggplot2`.

```{r exercise-viz, exercise=TRUE, exercise.setup="data2"}

```

```{r exercise-viz-hint-1, eval=FALSE}
boxplot(bloodpr ~ group, 
        data = C4T1,
        col = "bisque",
        ylab = "Blood Pressure", 
        xlab = "Treatment Group")
```

## The Omnibus ANOVA

### Running `aov()`

As we learned in the first tutorial, `aov()` requires a **formula** (`dependent ~ independent`) and the dataset. Let's run the omnibus one-way ANOVA:

```{r omnibus-setup, echo=FALSE}
library(dplyr)
library(AMCP)
library(emmeans)
library(ggplot2)

data(chapter_4_table_1)
C4T1 <- chapter_4_table_1

C4T1 <- C4T1 |> 
  mutate(group = case_match(cond,
    1 ~ "Drug Therapy",
    2 ~ "Biofeedback",
    3 ~ "Diet",
    4 ~ "Combination"
  )) |>
  mutate(group = factor(group), cond = factor(cond))
```

```{r omnibus1, exercise=TRUE, exercise.setup="omnibus-setup"}
mdl <- aov(bloodpr ~ group, data = C4T1)
summary(mdl)
```

###

The omnibus F-test tells us whether there is *any* significant difference among the four group means. If the p-value is below our alpha level, we reject the null hypothesis that all group means are equal.

But the omnibus test doesn't tell us **where** the differences are. For that, we need follow-up comparisons --- and this is where `emmeans` comes in.

```{r quiz-omnibus, echo=FALSE}
question_text(
  "Based on the omnibus ANOVA output above, is the overall F-test significant at alpha = .05? What does this tell us (and what does it NOT tell us)?",
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4
)
```

## Introduction to `emmeans`

### What is `emmeans`?

The `emmeans` package (Estimated Marginal Means) is a powerful and flexible tool for conducting follow-up comparisons after fitting a model. It can handle:

- **Pairwise comparisons** between all group means
- **Custom (complex) contrasts** that test specific hypotheses
- Adjustments for **multiple comparisons** (e.g., Tukey, Bonferroni)
- Comparisons in more complex designs (factorial, repeated measures --- we will cover these later)

The basic workflow is:

1. Fit a model with `aov()` (or `lm()`)
2. Extract the estimated marginal means with `emmeans()`
3. Conduct comparisons with `pairs()` or `contrast()`

###

To install `emmeans` (if you haven't already):

```{r, eval=FALSE}
install.packages("emmeans")
library(emmeans)
```

### Estimated Marginal Means

The first step after fitting a model is to extract the estimated marginal means (EMMs). These are the model-based estimates of each group mean.

The `emmeans()` function requires an already fitted model as its argument. In other words, we assign the `aov()` result to an object, then provide that object to `emmeans()`. It recognizes the original data and the variable names for the groups. 

To get the groups, we ask for part of the formula again, with ` ~ group_var_name` within in the function. 

```{r emmeans-setup, echo=FALSE}
library(dplyr)
library(AMCP)
library(emmeans)
library(ggplot2)

data(chapter_4_table_1)
C4T1 <- chapter_4_table_1

C4T1 <- C4T1 |> 
  mutate(group = case_match(cond,
    1 ~ "Drug Therapy",
    2 ~ "Biofeedback",
    3 ~ "Diet",
    4 ~ "Combination"
  )) |>
  mutate(group = factor(group), cond = factor(cond))

#mdl <- aov(bloodpr ~ group, data = C4T1)
```

```{r emm1, exercise=TRUE, exercise.setup="emmeans-setup"}
mdl <- aov(bloodpr ~ group, data = C4T1)
emm <- emmeans(mdl, ~ group)
emm
```

The output shows each group's estimated mean, the standard error, degrees of freedom, and a confidence interval. In a balanced one-way design, these estimated means are identical to the simple group means we computed with `summarize()`. The advantage of `emmeans` becomes clear in more complex or unbalanced designs, and --- as we'll see next --- for conducting comparisons.

## Pairwise Comparisons

### All Pairwise Tests

The simplest follow-up after a significant omnibus ANOVA is to compare every pair of group means. The `pairs()` function does this automatically:

```{r pairs1, exercise=TRUE, exercise.setup="emmeans-setup"}
mdl <- aov(bloodpr ~ group, data = C4T1)
emm <- emmeans(mdl, ~ group)
pairs(emm)
```

###

By default, `pairs()` applies a **Tukey adjustment** for multiple comparisons, which controls the familywise error rate across all pairwise tests. The output shows the estimated difference (`estimate`), standard error, degrees of freedom, t-ratio, and the adjusted p-value for each pair.

### Changing the Adjustment Method

You can change the p-value adjustment method with the `adjust` argument. Common options include `"tukey"` (default), `"bonferroni"`, `"sheffe"`, `"dunnettx"`, `"fdr"` (Benjamini-Hochberg), and `"none"`:

```{r pairs2, exercise=TRUE, exercise.setup="emmeans-setup"}
mdl <- aov(bloodpr ~ group, data = C4T1)
emm <- emmeans(mdl, ~ group)

# No adjustment (unadjusted p-values)
pairs(emm, adjust = "none")
```

###

```{r pairs3, exercise=TRUE, exercise.setup="emmeans-setup"}
mdl <- aov(bloodpr ~ group, data = C4T1)
emm <- emmeans(mdl, ~ group)

# Bonferroni adjustment
pairs(emm, adjust = "bonferroni")
```

### Exercise: Interpret Pairwise Comparisons

Run the pairwise comparisons with Tukey adjustment and answer the question below.

```{r exercise-pairs, exercise=TRUE, exercise.setup="emmeans-setup"}
mdl <- aov(bloodpr ~ group, data = C4T1)
emm <- emmeans(mdl, ~ group)
pairs(emm)
```

```{r quiz-pairs, echo=FALSE}
question_text(
  "Which pair(s) of groups show a statistically significant difference at alpha = .05 after Tukey adjustment? What does this tell us about the hypertension treatments?",
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4
)
```

## Custom (Complex) Contrasts

### Beyond Pairwise Comparisons

Pairwise comparisons test every possible pair, but sometimes we have specific, theory-driven hypotheses. For example:

- Does the **Combination** treatment differ from the average of the other three treatments?
- Do the **non-drug treatments** (Biofeedback, Diet, Combination) differ from **Drug Therapy** alone?

These are **complex contrasts** --- comparisons that involve more than two groups at a time. The `emmeans` package handles these with the `contrast()` function.

### Setting Up Custom Contrasts

To define a custom contrast, we create a named *list* of coefficient vectors. Each vector must have one value per group, and the coefficients must sum to zero (the definition of a contrast).

First, let's check the order of the groups in our `emmeans` object, since the coefficients must match this order:

```{r custom-setup, echo=FALSE}
library(dplyr)
library(AMCP)
library(emmeans)

data(chapter_4_table_1)
C4T1 <- chapter_4_table_1

C4T1 <- C4T1 |> 
  mutate(group = case_match(cond,
    1 ~ "Drug Therapy",
    2 ~ "Biofeedback",
    3 ~ "Diet",
    4 ~ "Combination"
  )) |>
  mutate(group = factor(group), cond = factor(cond))

mdl <- aov(bloodpr ~ group, data = C4T1)
emm <- emmeans(mdl, ~ group)
```

```{r custom1, exercise=TRUE, exercise.setup="custom-setup"}
mdl <- aov(bloodpr ~ group, data = C4T1)
emm <- emmeans(mdl, ~ group)
emm
```

###

The groups are listed in the output. By default, R orders factor levels alphabetically, so the order here is: Biofeedback, Combination, Diet, Drug Therapy. **This is the order your coefficients must follow.**

Now let's test the hypothesis that the Combination treatment differs from the average of the other three. The contrast coefficients for this comparison are:

| Group | Biofeedback | Combination | Diet | Drug Therapy |
|-------|:-----------:|:-----------:|:----:|:------------:|
| Coefficient | 1/3 | -1 | 1/3 | 1/3 |

These sum to zero (1/3 + (-1) + 1/3 + 1/3 = 0), confirming this is a valid contrast.

```{r custom2, exercise=TRUE, exercise.setup="custom-setup"}
emm <- emmeans(mdl, ~ group)

# Combination vs. average of all others
combo_vs_rest <- list("Combo vs Others" = c(1/3, -1, 1/3, 1/3))

contrast(emm, combo_vs_rest)
```

### Testing Multiple Custom Contrasts

We can test multiple contrasts at once by adding more entries to the list. Let's also test whether Drug Therapies (Drug + Combo) differ from the average of the non-drug treatments (Biofeedback, Diet):

```{r custom3, exercise=TRUE, exercise.setup="custom-setup"}
emm <- emmeans(mdl, ~ group)

my_contrasts <- list(
  "Combo vs Others"       = c(1/3, -1, 1/3, 1/3),
  "Drug vs Non-Drug"      = c(-1/2, 1/2, -1/2, 1/2)
)

contrast(emm, my_contrasts)
```

###

You can also apply a p-value adjustment across your custom contrasts:

```{r custom4, exercise=TRUE, exercise.setup="custom-setup"}
emm <- emmeans(mdl, ~ group)

my_contrasts <- list(
  "Combo vs Others"       = c(1/3, -1, 1/3, 1/3),
  "Drug vs Non-Drug"      = c(-1/2, 1/2, -1/2, 1/2)
)

contrast(emm, my_contrasts, adjust = "bonferroni")
```

This is helpful in case we have very specific pairwise comparisons, and we don't want the multiple testing penalty of testing **all** pairs. 

### Confidence Intervals for Contrasts

You can obtain confidence intervals for any comparison by wrapping the result in `confint()`:

```{r custom5, exercise=TRUE, exercise.setup="custom-setup"}
emm <- emmeans(mdl, ~ group)

combo_vs_rest <- list("Combo vs Others" = c(1/3, -1, 1/3, 1/3))

confint(contrast(emm, combo_vs_rest))
```

### Effect Sizes

The `emmeans` package can also provide standardized effect sizes for contrasts using `eff_size()`. This computes a Cohen's *d*-type measure using the pooled within-group SD:

```{r custom6, exercise=TRUE, exercise.setup="custom-setup"}
emm <- emmeans(mdl, ~ group)

# Effect sizes for pairwise comparisons
eff_size(emm, sigma = sigma(mdl), edf = df.residual(mdl))
```

The `sigma(mdl)` extracts the residual standard deviation (i.e., the square root of $MS_W$), and `df.residual(mdl)` provides the error degrees of freedom.

## Exercise: Custom Contrasts

### Your Turn

Using the `chapter_4_table_1` data (already loaded as `C4T1`), set up and test a custom contrast comparing **Biofeedback vs. Drug Therapy**. Remember, the group order is alphabetical: Biofeedback, Combination, Diet, Drug Therapy.

```{r exercise-contrast, exercise=TRUE, exercise.setup="custom-setup"}
emm <- emmeans(mdl, ~ group)

# Define the contrast: Biofeedback vs Drug Therapy
my_contrast <- list(
  "Bio vs Drug" = c(___, ___, ___, ___)
)

contrast(emm, my_contrast)
```

```{r exercise-contrast-hint-1, eval=FALSE}
# Biofeedback gets +1, Drug Therapy gets -1, others get 0
my_contrast <- list(
  "Bio vs Drug" = c(1, 0, 0, -1)
)
```

```{r exercise-contrast-solution}
emm <- emmeans(mdl, ~ group)

my_contrast <- list(
  "Bio vs Drug" = c(1, 0, 0, -1)
)

contrast(emm, my_contrast)
```

###

Now try something more complex. Test whether the average of the two "single-method" treatments (Drug Therapy and Diet) differs from the average of the two "alternative" treatments (Biofeedback and Combination).

```{r exercise-contrast2, exercise=TRUE, exercise.setup="custom-setup"}
emm <- emmeans(mdl, ~ group)

# Your contrast here


```

```{r exercise-contrast2-hint-1, eval=FALSE}
# Biofeedback = +1/2, Combination = +1/2, Diet = -1/2, Drug Therapy = -1/2
my_contrast <- list(
  "Alt vs Traditional" = c(1/2, 1/2, -1/2, -1/2)
)

contrast(emm, my_contrast)
```

## Knowledge Check

```{r quiz-final, echo=FALSE}
quiz(
  caption = "Test Your Understanding",
  
  question("What does the omnibus ANOVA F-test tell us?",
    answer("Which specific groups differ from each other"),
    answer("Whether there is any significant difference among the group means", correct = TRUE),
    answer("The effect size of the treatment"),
    answer("Whether assumptions are met"),
    allow_retry = TRUE
  ),
  
  question("In emmeans, what function is used to conduct all pairwise comparisons?",
    answer("contrast()"),
    answer("emmeans()"),
    answer("pairs()", correct = TRUE),
    answer("pairwise()"),
    allow_retry = TRUE
  ),
  
  question("What must be true about the coefficients in a contrast vector?",
    answer("They must all be positive"),
    answer("They must all be integers"),
    answer("They must sum to zero", correct = TRUE),
    answer("They must sum to one"),
    allow_retry = TRUE
  ),
  
  question("Why is the order of groups in the emmeans output important when specifying custom contrasts?",
    answer("It determines the p-value"),
    answer("The contrast coefficients must match the group order so each weight is applied to the correct group mean", correct = TRUE),
    answer("It affects the degrees of freedom"),
    answer("The order doesn't matter"),
    allow_retry = TRUE
  ),
  
  question("What adjustment does pairs() apply by default to control for multiple comparisons?",
    answer("Bonferroni"),
    answer("None"),
    answer("Holm"),
    answer("Tukey", correct = TRUE),
    allow_retry = TRUE
  )
)
```

## Summary

### What We Covered

In this tutorial, you learned:

- How to **recode variables** using `case_match()` within `mutate()`
- The **native pipe** `|>` as an alternative to `%>%` from the tidyverse
- How to run and interpret a one-way **omnibus ANOVA** with `aov()`
- How to extract **estimated marginal means** with `emmeans()`
- How to conduct **pairwise comparisons** with `pairs()` and adjust for multiple testing
- How to define and test **custom (complex) contrasts** with `contrast()`
- How to obtain **confidence intervals** and **effect sizes** for comparisons

### The `emmeans` Workflow

The general workflow for follow-up tests is:

1. `mdl <- aov(DV ~ IV, data = mydata)` --- fit the model
2. `emm <- emmeans(mdl, ~ IV)` --- extract estimated marginal means
3. `pairs(emm)` --- for pairwise comparisons, **or**
4. `contrast(emm, my_list)` --- for custom contrasts

We will continue to use `emmeans` in future tutorials as we move to factorial and repeated measures designs.
